Knowledge Representation for Service Robots: Supply and Demand

### 2012-05-14 krs-survey.tex
Filled section on Explanation

\subsubsection{Open World and Close World Assumptions}

@@

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf C. Reasoning}
            child {node [taxon] {{\bf C.10}. Learning}}
            child {node [taxon] {{\bf C.9}. Naive physics}}
            child {node [taxon] {{\bf C.8}. Planning}}
            child {node [taxon] {{\bf C.7}. Prediction}}
            child {node [taxon] {{\bf C.6}. Presupposition}}
            child {node [taxon] {{\bf C.5}. Non-monotonicity}}
            child {node [taxon] {{\bf C.4}. Uncertainty}}
            child {node [taxon] {{\bf C.3}. Lazy evaluation}}
            child {node [taxon] {{\bf C.2}. Instantiation and structural alteration}}
            child {node [taxon] {{\bf C.1}. Standard reasoning}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

@@

\subsubsection{Prediction, projection and diagnosis tasks}
\label{sect|prediction-projection}

@@

>>>>

\subsubsection{Open World and Close World Assumptions}
\label{sect|owa-cwa}

@@

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf C. Reasoning}
            child {node [taxon] {{\bf C.10}. Learning}}
            child {node [taxon] {{\bf C.9}. Naive physics}}
            child {node [taxon] {{\bf C.8}. Planning}}
            child {node [taxon] {{\bf C.7}. Prediction and explanation}}
            child {node [taxon] {{\bf C.6}. Presupposition}}
            child {node [taxon] {{\bf C.5}. Non-monotonicity}}
            child {node [taxon] {{\bf C.4}. Uncertainty}}
            child {node [taxon] {{\bf C.3}. Lazy evaluation}}
            child {node [taxon] {{\bf C.2}. Instantiation and structural alteration}}
            child {node [taxon] {{\bf C.1}. Standard reasoning}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

@@

\subsubsection{Prediction, projection, explanation}
\label{sect|prediction-projection}

@@

\paragraph{Explanation} Diagnosis is also linked to the \emph{explanation} or
\emph{justification} capabilities that may be offered by the knowledge
representation system. The explanation of an entailment is the sequence (or
set of sequences if several are possible) of reasoning steps that allow to
conclude. An explanation can also conversely explain why a statement leads to a
contradiction.

The following example shows an explanation for an inconsistency in a particular
knowledge base: by adding the statement \stmt{robot1 belongsTo human1}, we
observe that an inconsistency is triggered. The reasoner provides the four
following observations\footnote{This example uses the \emph{Manchester syntax},
\url{http://www.w3.org/TR/owl2-manchester-syntax/}} to explain the
inconsistency:

\begin{quote}
\scriptsize
\begin{enumerate}
    \item \concept{myself} {\bf Type} \concept{belongsTo} {\bf some} \concept{Thing}
    \item \concept{belongsTo} {\bf Domain} \concept{Artifact}
    \item \concept{myself} {\bf Type} \concept{Agent}
    \item {\bf DisjointClasses:} \concept{Agent}, \concept{Artifact}
\end{enumerate}
\normalsize
\end{quote}

This explanation directly portrays the underlying structure of the knowledge
model. A KRS may provide mechanisms to expose this kind of analysis, either
automatically when inconsistencies survene, or on demand.

Explanation of contradictions plays a particular role for robots: from a
cognitive point of view, a logical inconsistency (\ie, a contradiction) can be
viewed as a \emph{cognitive conflict} or \emph{cognitive dissonance} (\ie, two
incompatible models of the world that must be dealt with). Being able to expose
and explain such cognitive conflicts eases the control of the behaviour of the
robot in unexpected semantic situations, and forms a first step towards an
adquate reaction.

Cognitive dissonance is also identified by developmental psychologists like
Piaget as a motivation factor for a child to progress through the various
stages of its cognitive development. This has been also studied in
robotics~\cite{Oudeyer2007}.

The cognitive capability of \emph{justification} is also intimately linked to
the meta-cognition capability and participates to the overall \emph{cognitive
observability} of the system.
